---
title: "Pupillometry and AX-CPT"
author: "Jeremy Elman"
date: "`r Sys.Date()`"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

Load libraries
```{r results='hide', message=FALSE, warning=FALSE}
library(ggplot2)
library(gridExtra)
library(gtools)
library(lme4)
library(arm)
library(lmerTest)
library(dplyr)
library(magrittr)
library(sjPlot)
library(sjmisc)
library(multcomp)
source("K:/code/summarySE.R")
source("K:/code/normDataWithin.R")
source("K:/code/summarySEwithin.R")
sjp.setTheme(base=theme_bw())
```

Load data.
```{r}
pupilAXCPT = read.csv("K:/AX-CPT/data/pupil_AX-CPT.csv", 
                         stringsAsFactors = FALSE)
pupilAXCPT = pupilAXCPT[!is.na(pupilAXCPT$PCA) & !is.na(pupilAXCPT$rMCI_cons_v2),]
pupilAXCPT = subset(pupilAXCPT, vetsa2=='vetsa2')
str(pupilAXCPT)
```

Recode two-level factors to 0/1. Center continuous data.
*TRANSFORM*
```{r}
# Create factors
pupilAXCPT$facLoad = factor(pupilAXCPT$Load)
pupilAXCPT$rMCI_cons_v2 = factor(pupilAXCPT$rMCI_cons_v2)
contr.backward.diff = matrix(c(-2/3, 1/3, 1/3,-1/3,-1/3,2/3), ncol = 2)
contrasts(pupilAXCPT$facLoad) = contr.backward.diff
pupilAXCPT$MZ = ifelse(pupilAXCPT$zyg14==1, 1, 0)
pupilAXCPT$site_v2rev = ifelse(pupilAXCPT$site_v2rev==1, 1, 0)
pupilAXCPT$Device = factor(pupilAXCPT$LR_Device)
pupilAXCPT$acamedbin = ifelse(pupilAXCPT$acamedtot==0, 0, 1)

#Bin MCI data into 0 = No impairment, 1 = Single domain MCI, 2 = Multi-domain MCI.
pupilAXCPT$rMCI_cons_bin = ifelse(pupilAXCPT$rMCI_cons_v2==0, 0, 
                                        ifelse(pupilAXCPT$rMCI_cons_v2==1 | 
                                               pupilAXCPT$rMCI_cons_v2==2, 1, 2))
pupilAXCPT$rMCI_cons_bin = factor(pupilAXCPT$rMCI_cons_bin)

# Center continuous variables to interpret intercept at mean value (rather than 0).
# Subject specific variables centered at subject level, trial specific variables centered at trial level.
contTrialVars = c("PCA","pctPCA","adjPCA")
contSubjVars = c("age_v2","afqtpcttran_v2","DSFMAX_v2","dprime")

# Center trial specific variables
for (x in contTrialVars) {
  newcol = paste0('c.',x)
  pupilAXCPT[[newcol]] = as.numeric(scale(pupilAXCPT[[x]], center=TRUE, scale=FALSE))
}

# Center subject specific variables
subjDF = pupilAXCPT[c("vetsaid",contSubjVars)]
subjDF %<>% group_by(vetsaid) %>% dplyr::summarise_each(funs(mean))
nums <- sapply(subjDF, is.numeric)
c.subjDF = as.data.frame(apply(subjDF[,nums],2, function(y) y - mean(y, na.rm=TRUE)))
names(c.subjDF) = paste0("c.",names(c.subjDF))
c.subjDF$vetsaid = subjDF$vetsaid
pupilAXCPT = left_join(pupilAXCPT, c.subjDF, by="vetsaid")

# Log transform BY RT measures
pupilAXCPT$log.bymeanrt = log(pupilAXCPT$bymeanrt)
pupilAXCPT$log.bymedianrt = log(pupilAXCPT$bymedianrt)
pupilAXCPT$log.bystdrt = log(pupilAXCPT$bystdrt)
pupilAXCPT$log.bycvrt = log(pupilAXCPT$bycvrt)

# Get BY RT standard deviation with mean RT regressed out
pupilAXCPT %<>%
  dplyr::select(vetsaid, bymeanrt, bystdrt) %>%
  group_by(vetsaid) %>%
  dplyr::summarise(bymeanrt = first(bymeanrt),
                   bystdrt = first(bystdrt)) %>%
  do(data.frame(vetsaid = .[["vetsaid"]],
                resid.bystdrt = resid(lm(bystdrt ~ bymeanrt, data=.)))) %>%
  inner_join(pupilAXCPT, by="vetsaid")

# Create quantile groups of log RT data
pupilAXCPT %<>%
  group_by(vetsaid) %>%
  dplyr::summarise(log.bymeanrt = first(log.bymeanrt),
                   log.bymedianrt = first(log.bymedianrt),
                   log.bystdrt = first(log.bystdrt),
                   log.bycvrt = first(log.bycvrt),
                   resid.bystdrt = first(resid.bystdrt)) %>%
  mutate(q.log.bymeanrt = quantcut(log.bymeanrt, q=4, labels=seq(1,4)),
         q.log.bymedianrt = quantcut(log.bymedianrt, q=4, labels=seq(1,4)),
         q.log.bystdrt = quantcut(log.bystdrt, q=4, labels=seq(1,4)),
         q.log.bycvrt = quantcut(log.bycvrt, q=4, labels=seq(1,4)),
         q.resid.bystdrt = quantcut(resid.bystdrt, q=4, labels=seq(1,4))) %>%
  dplyr::select(vetsaid,q.log.bymeanrt,q.log.bymedianrt,q.log.bystdrt,
                q.log.bycvrt,q.resid.bystdrt) %>%
  inner_join(pupilAXCPT, by="vetsaid")

# Create residPCA by regressing BaselineDiameter from PCA and taking the residuals.
lme.PCA.resid = lmer(PCA ~ BaselineDiameter + (1 | case/vetsaid),data=pupilAXCPT)
pupilAXCPT$residPCA = resid(lme.PCA.resid)
```

# Basic sample descriptives
```{r, include=FALSE}
#Create dataset with subject level variables to assess relationships with BOLD variance.

subjDatAXCPT = pupilAXCPT %>% 
            dplyr::select(-contains("PCA"),-contains("Load")) %>%
            group_by(vetsaid) %>%
            summarise_each(funs(first))
```

How many subjects?
```{r}
n_distinct(subjDatAXCPT$vetsaid)
```

How many attrition replacements?
```{r}
dplyr::count(subjDatAXCPT, vetsa2)
```

How many twin pairs vs unpaired twins? 
```{r}
subjDatAXCPT %>%
  group_by(case) %>%
  dplyr::summarise(n_twins = n_distinct(vetsaid)) %>%
  dplyr::count(n_twins)
```


How many MZ and DZ pairs? (excludes unpaired subjects)
```{r}
subjDatAXCPT %>% 
  group_by(case) %>%
  mutate(n_twins = n_distinct(vetsaid)) %>%
  filter(n_twins > 1) %>%
  dplyr::summarise(zyg = mean(zyg14)) %>%
  group_by(zyg) %>%
  dplyr::count(zyg)
```

How many levels of load did subjects complete? Shows number of subjects who completed 1, 2, or 3 levels (corresponding to digit spans of 3, 6, and 9).
```{r}
pupilAXCPT %>%
  group_by(vetsaid) %>%
  dplyr::summarise(load = n()) %>%
  dplyr::count(load)
```

How many of each MCI group?
```{r}
dplyr::count(subjDatAXCPT, rMCI_cons_v2)
```

How many of each MCI group binned by single or multi-doman MCI?
```{r}
dplyr::count(subjDatAXCPT, rMCI_cons_bin)
```

Look at frequencies of MCI measures
```{r}
mci.p1 = sjp.frq(subjDatAXCPT$rMCI_cons_v2, 
          title="MCI groups",
          axisLabels.x=c("Normal","Single Domain\nNon-amnestic","Single Domain\nAmnestic"),
          axisLimits.y = c(0,850))

mci.p2 = sjp.frq(subjDatAXCPT$rMCI_cons_bin, 
          title="Binned MCI groups",
          axisLabels.x=c("Normal","Single Domain"),
          axisLimits.y = c(0,850))

grid.arrange(mci.p1$plot,mci.p2$plot)
```

```{r, include=FALSE}
pdf("K:/AX-CPT/results/MCI_groups.pdf", width=8, height=6)
grid.arrange(mci.p1$plot,mci.p2$plot)
dev.off()
```


Plot pctPCA and residPCA data to check for ceiling effects on pupil dilation
```{r}
pctPcaPlot = ggplot(pupilAXCPT, aes(x=pctPCA)) + 
  geom_histogram(fill="steel blue",color="black") + 
  facet_wrap(~ Load, nrow=3) + 
  ggtitle("Histogram of pctPCA by Load") + 
  xlab("% PCA") + theme_bw() 

residPcaPlot = ggplot(pupilAXCPT, aes(x=residPCA)) + 
  geom_histogram(fill="steel blue",color="black") + 
  facet_wrap(~ Load, nrow=3) + 
  ggtitle("Histogram of residPCA by Load") + 
  xlab("PCA (adjusted for baseline)") + theme_bw() 

grid.arrange(pctPcaPlot, residPcaPlot, ncol=2)
```

```{r, include=FALSE}
pdf("K:/AX-CPT/results/PCAxLoad.pdf")
grid.arrange(pctPcaPlot, residPcaPlot, ncol=2)
dev.off()
```



# AX-CPT performance 

## Error Rates

d' by MCI group
```{r}
dprime.p1 = ggplot(subjDatAXCPT, aes(x=dprime)) + 
  geom_histogram(fill="steel blue",color="black") + 
  theme(plot.title=element_text(face="bold")) + theme_bw(14)

dprime.p2 = ggplot(subjDatAXCPT, aes(x=rMCI_cons_bin,y=dprime)) + 
              geom_boxplot(fill="firebrick",alpha=.75) +
              geom_jitter(alpha=0.3,position=position_jitter(width = 0.1)) +
              theme(plot.title=element_text(face="bold")) +
              theme_bw(14)
grid.arrange(dprime.p1,dprime.p2, ncol=2, top="d' by MCI status")

summary(lm(dprime ~ rMCI_cons_bin, data=subjDatAXCPT))
```

```{r,include=FALSE}
pdf("K:/AX-CPT/results/dprime_X_MCI.pdf")
grid.arrange(dprime.p1,dprime.p2, ncol=2, top="d' by MCI status")
dev.off()
```

AX False alarm rates
```{r}
axfa.p1 = ggplot(subjDatAXCPT, aes(x=axfarate)) + 
  geom_histogram(fill="steel blue",color="black") + 
  theme(plot.title=element_text(face="bold")) + theme_bw(14)

axfa.p2 = ggplot(subjDatAXCPT, aes(x=rMCI_cons_bin,y=axfarate)) + 
              geom_boxplot(fill="firebrick",alpha=.75) +
              geom_jitter(alpha=0.3,position=position_jitter(width = 0.1)) +
              theme(plot.title=element_text(face="bold")) +
              theme_bw(14)
grid.arrange(axfa.p1, axfa.p2, ncol=2, top="AX FA Rate by MCI status")
summary(lm(axfarate ~ rMCI_cons_bin, data=subjDatAXCPT))
```

```{r,include=FALSE}
pdf("K:/AX-CPT/results/axfarate_X_MCI.pdf")
grid.arrange(axfa.p1, axfa.p2, ncol=2, top="AX FA Rate by MCI status")
dev.off()
```


BX False alarm rates
```{r}
bxfa.p1 = ggplot(subjDatAXCPT, aes(x=bxfarate)) + 
  geom_histogram(fill="steel blue",color="black") + 
  theme(plot.title=element_text(face="bold")) + theme_bw(14)

bxfa.p2 = ggplot(subjDatAXCPT, aes(x=rMCI_cons_bin,y=bxfarate)) + 
              geom_boxplot(fill="firebrick",alpha=.75) +
              geom_jitter(alpha=0.3,position=position_jitter(width = 0.1)) +
              theme(plot.title=element_text(face="bold")) +
              theme_bw(14)
grid.arrange(bxfa.p1, bxfa.p2, ncol=2, top="BX FA Rate by MCI status")
summary(lm(bxfarate ~ rMCI_cons_bin, data=subjDatAXCPT))
```

```{r,include=FALSE}
pdf("K:/AX-CPT/results/bxfarate_X_MCI.pdf")
grid.arrange(bxfa.p1, bxfa.p2, ncol=2, top="BX FA Rate by MCI status")
dev.off()
```

AY False alarm rates
```{r}
ayfa.p1 = ggplot(subjDatAXCPT, aes(x=ayfarate)) + 
  geom_histogram(fill="steel blue",color="black") + 
  theme(plot.title=element_text(face="bold")) + theme_bw(14)

ayfa.p2 = ggplot(subjDatAXCPT, aes(x=rMCI_cons_bin,y=ayfarate)) + 
              geom_boxplot(fill="firebrick",alpha=.75) +
              geom_jitter(alpha=0.3,position=position_jitter(width = 0.1)) +
              theme(plot.title=element_text(face="bold")) +
              theme_bw(14)
grid.arrange(ayfa.p1, ayfa.p2, ncol=2, top="AY FA Rate by MCI status")
summary(lm(ayfarate ~ rMCI_cons_bin, data=subjDatAXCPT))
```

```{r,include=FALSE}
pdf("K:/AX-CPT/results/ayfarate_X_MCI.pdf")
grid.arrange(ayfa.p1, ayfa.p2, ncol=2, top="AY FA Rate by MCI status")
dev.off()
```

## Reaction Times

BY mean reaction time
```{r}
bymeanrt.p1 = ggplot(subjDatAXCPT, aes(x=by)) + 
  geom_histogram(fill="steel blue",color="black") + 
  theme(plot.title=element_text(face="bold")) + theme_bw(14)

ayfa.p2 = ggplot(subjDatAXCPT, aes(x=rMCI_cons_bin,y=ayfarate)) + 
              geom_boxplot(fill="firebrick",alpha=.75) +
              geom_jitter(alpha=0.3,position=position_jitter(width = 0.1)) +
              theme(plot.title=element_text(face="bold")) +
              theme_bw(14)
grid.arrange(ayfa.p1, ayfa.p2, ncol=2, top="AY FA Rate by MCI status")
summary(lm(ayfarate ~ rMCI_cons_bin, data=subjDatAXCPT))
```

```{r,include=FALSE}
pdf("K:/AX-CPT/results/ayfarate_X_MCI.pdf")
grid.arrange(ayfa.p1, ayfa.p2, ncol=2, top="AY FA Rate by MCI status")
dev.off()
```